SecFFT: Safeguarding Federated Fine-tuning for Large Vision Language Models against Stealthy Backdoor Attacks in IoRT Networks

### 2024-09-13 main-v2.tex
docs: 理论(Let)+en(zhou)

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{figures/fig1.PNG}
\caption{FFL Illustration in Heterogeneous UAV Networks: Taking target recognition as an example, UAV node group patrolling in urban areas may collect a higher proportion of ``automobile'' samples (e.g., UN1 and UN2), while UAV groups patrolling in port areas will acquire more ``ship'' samples (e.g., UN6 and UN7). Consequently, the dataset distributions across the entire network often exhibit significant imbalance, while simultaneously displaying a certain degree of similarity within localized regions.
}
\label{fig1:UAV-Network}
\end{figure}

@@

\section{System model} %\& Preliminaries}
%%symbols 表格
\begin{table}[!t]
{
\renewcommand{\arraystretch}{1.2}
\caption{Summary of Main Notation}
\label{notations}
\vspace{0em}
\centering
\begin{tabular}{| c || l |}
\hline
\textbf{Symbol} & \multicolumn{1}{c |}{\textbf{Description}}\\
    \hline
    $U_i,\mathcal{D}_i$ & The $i$-th UAV devices and its dataset\\
    \hline
    $n^{s_{m}, i},\mathcal{P}^i$ & The size and normalized distribution of $\mathcal{D}_i$\\
    \hline
    $\mathcal{P}^{Tar}, w$ & Target distribution and model of learning task $Job$\\
    \hline
    $T$ & The maximum communication rounds of $Job$\\
    \hline
    $C,K$ & The set and number of all peer communities\\
    \hline
    $S,M$ & The set and number of all colleague communities\\
    \hline
    %$L$ & The number of clients in each super node\\
    %\hline
    $c_k$ & The $k$-th peer community\\
    \hline
    $s_m$ & The $m$-th colleague community\\ 
    \hline
    $\Delta \widetilde{\omega}_{t}^{s_{m},i}$ & The weight updates from $U_i \in s_m$ in round $t$\\
    \hline
    $p^{s_{m}, i}$ & The aggregation coefficient for $U_i$ within $s_m$\\
    \hline
    $\Delta\widetilde{\omega}_{t}^{s_{m}}$ & The weight updates from $s_m$ in round $t$\\
    \hline
    $p^{s_{m}}$ & The aggregation coefficient for $\Delta\widetilde{\omega}_{t}^{s_{m}}$ from $s_m$\\
    \hline
    $E_k$ & The clipping bound of peer community $c_k$\\
    \hline
    $Sen_k$ & The maximum sensitivity of $c_k$\\
    \hline
    $\varepsilon_k,\delta_k$& The parameters of differential privacy \\
    \hline
    $L$& The minimum size of colleague community\\
    \hline
    $\alpha^{a,b}_{t}$& The cosine similarity between $a$ and $b$\\
    \hline
    $PI^{c_k \rightarrow i}_{t}$ & The positive interaction between $c_k$ and $U_i$\\
    \hline
    $NI^{c_k \rightarrow i}_{t}$& The positive interaction between $c_k$ and $U_i$\\
    \hline
    $\Upsilon^{c_k \rightarrow i}_{t}$& The peer reputation opinion vector of $c_k$\\
    \hline
    $\Phi^{c_k \rightarrow i}_{t}$& $c_k$'s credit evaluation towards $U_i$ in round $t$\\
    \hline
    $\Upsilon^{s_m}_{t}$ & The colleague reputation opinion vector towards $s_m$\\
    \hline
    $\Phi^{s_m \rightarrow i}_{t}$& The colleague credit evaluation for $\forall U_i \in s_m$\\
    \hline
    $\Upsilon^{i}_{t},\Phi^{i}_{t}$& The comprehensive reputation and credit of $U_i$\\
    \hline
    $\psi^{i}_t$& The fairness-guaranteed selection probability\\
    \hline
    \end{tabular}
        }
     \vspace{-1.5em}
     \end{table}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{figures/fig2.PNG}
\caption{Stratified community construction in a heterogeneous environment.}
\label{fig2:framework}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{figures/fig3.PNG}
\caption{The workflow of MR-FFL.}
\label{fig3:workflow}
\end{figure}

@@

@@

\bibliographystyle{IEEEtran}
\bibliography{ref}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{figures/fig4-history.jpg}
\caption{Combining multiple rounds of historical identification to identify malicious nodes
}
\label{fig4:history}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{img/SecFFT-场景.pdf}
\caption{Combining multiple rounds of historical identification to identify malicious nodes
}
\label{fig4:history}
\end{figure}

>>>>

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/fig1.pdf}
    \caption{Benign and Malicious Nodes Before Aggregation}
    \label{fig1:scene}
\end{figure}

\section{Related Works}
%后门攻击对深度学习模型构成了严重的安全威胁，它旨在通过对模型输入添加特殊扰动来诱导目标模型产生不当行为，例如图像分类器中的错误分类等。我们将现有针对模型的后门攻击分为两种，即空间隐蔽性后门攻击和时间隐蔽性后门攻击。
Backdoor attacks pose a significant security threat to deep learning models, aiming to induce improper behavior in the target model by adding specific perturbations to the input, such as misclassification in image classifiers. We classify existing backdoor attacks on models into two types: spatially covert backdoor attacks and temporally covert backdoor attacks.

%空间隐蔽性后门攻击是指攻击者通过微调模型参数、修改输入特征或操纵模型训练过程等方式，使得现有防御方法无法在特征空间中有效区分恶意攻击者与良性参与者，从而有效逃避检测。Shadowcast\cite{xu2024shadowcast}通过在特征空间中引入微小、不可察觉的扰动，操纵视觉语言模型（VLMs）在特定视觉输入下生成误导性的文本输出。BadClip\cite{liang2024badclip}提出了一种针对多模态对比学习模型（如CLIP）的双嵌入引导后门攻击方法，通过优化视觉触发模式，使其在嵌入空间中接近目标文本语义，从而在不显著改变模型参数的情况下，植入难以被检测到的后门。Flip\cite{jha2023label}通过修改训练数据的标签，使得模型训练轨迹接近于带有后门的专家模型，从而在特征空间中达到隐蔽攻击的效果，即使在图像干净的情况下也能逃避现有检测方法。
Spatially Covert Backdoor Attacks:
Spatially covert backdoor attacks refer to attacks where adversaries fine-tune model parameters, modify input features, or manipulate the model training process in a way that existing defense methods cannot effectively distinguish between malicious attackers and benign participants in the feature space, thereby evading detection. Shadowcast \cite{xu2024shadowcast} manipulates vision-language models (VLMs) to produce misleading textual outputs for specific visual inputs by introducing subtle, imperceptible perturbations in the feature space. BadClip \cite{liang2024badclip} proposes a dual embedding-guided backdoor attack method for multimodal contrastive learning models (e.g., CLIP), optimizing visual trigger patterns to make them closer to the target text semantics in the embedding space, thus implanting a hard-to-detect backdoor without significantly altering the model parameters. Flip \cite{jha2023label} modifies the labels of training data so that the model's training trajectory approximates that of an expert model with a backdoor, achieving covert attacks in the feature space that can evade existing detection methods even when the images are clean.

%时间隐蔽性后门攻击是指攻击者通过调整模型训练过程或优化后门植入等方式，使后门在恶意攻击者结束攻击后仍能长期存在，不被后续良性更新所覆盖，从而持续影响模型行为。 Chameleon\cite{dai2023chameleon}关注如何通过对比学习来调整中毒样本、干扰样本以及促进样本嵌入表示之间的关系，从而提高后门攻击的空间隐蔽性。ImgTrojan\cite{tao2024imgtrojan}通过在视觉语言模型的训练数据中注入少量恶意样本，使得这些后门即使在模型后续的良性训练中仍能保持有效，从而持续影响模型的行为。\cite{gu2023gradient}提出了一种新的更新控制方法，通过跨层更新幅度归一化和层内更新方向投影，解决了参数高效调优（PET）过程中后门攻击的遗忘问题，从而维持攻击的持久性和隐蔽性。A3FL\cite{zhang2024a3fl}通过对抗性适应损失函数来优化触发器的方式，使得后门能够在全局训练动态中持久存在，从而在联邦学习模型中实现高效且持久的攻击效果。
Temporally Covert Backdoor Attacks:
Temporally covert backdoor attacks involve adjusting the model training process or optimizing backdoor implantation such that the backdoor persists even after the malicious attacker has stopped the attack, remaining undetected by subsequent benign updates, and continuing to affect the model’s behavior. Chameleon \cite{dai2023chameleon} focuses on enhancing the spatial covert nature of backdoor attacks by adjusting the relationships between poisoned samples, interference samples, and promoting samples in contrastive learning. ImgTrojan \cite{tao2024imgtrojan} injects a small number of malicious samples into the training data of vision-language models, ensuring that these backdoors remain effective even after subsequent benign training, thereby continuously affecting the model's behavior. \cite{gu2023gradient} proposes a novel update control method that normalizes cross-layer update magnitudes and projects intra-layer update directions to address the issue of backdoor forgetting during parameter-efficient tuning (PET), thus maintaining the persistence and concealment of the attack. A3FL \cite{zhang2024a3fl} optimizes the trigger by using an adversarial adaptation loss function, enabling the backdoor to persist throughout global training dynamics, thereby achieving efficient and persistent attacks in federated learning models.
% 总之，现有的后门攻击策略在空间隐蔽性和时间隐蔽性方面都有了显著的进展。空间隐蔽性后门攻击通过优化特征空间中的触发器或更新，使得恶意行为难以被现有的检测方法发现，从而实现了更高效、更难以检测的攻击。而时间隐蔽性后门攻击则专注于提高后门的持久性，使得即使在恶意攻击结束后，这些后门依然能够在模型中长期存在并继续影响其行为。通过这些攻击方式的结合，攻击者能够在各种应用场景中有效地躲避检测，并确保攻击效果的持续性，这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。

%总的来看，现有的后门攻击方法已展现出显著的隐蔽性和持久性，能够有效规避当前的检测机制。这些攻击不仅在特征空间中巧妙隐藏恶意意图，还能在长期训练过程中保持攻击效果，对模型行为产生持续而深远的影响，也对模型安全构成了严峻挑战。这对现有的防御机制提出了更高的挑战，强调了未来在开发更加鲁棒的防御方法方面的迫切需求。
Overall, existing backdoor attack methods have demonstrated significant advances in both spatial and temporal concealment, effectively circumventing current detection mechanisms. These attacks not only cleverly hide malicious intent in the feature space but also maintain their effects throughout long-term training, posing a sustained and profound impact on model behavior, thus presenting a severe challenge to model security. This underscores the need for future research to focus on developing more robust defense methods.

\subsection{Defenses}
%在当前深度学习模型的安全性研究中，针对后门攻击的防御措施主要可以分为两大类：后门检测和后门缓解。% Backdoor Mitigation or Elimination
In the current research on the security of deep learning models, defense measures against backdoor attacks can be broadly categorized into two major types: backdoor detection and backdoor mitigation.

%后门检测主要旨在识别模型中已存在的后门攻击。在这一类中，SEER\cite{zhu2024seer}提出了一种用于视觉-语言模型的后门检测算法，通过在图像和文本模态的特征空间中联合搜索目标文本和图像触发器，成功实现了在多种场景下的后门检测。ASSET\cite{pan2023asset}提出了一种通过主动诱导后门样本和干净样本在模型行为上的差异，从而实现跨多种深度学习范式的鲁棒后门数据检测方法，并在端到端监督学习、自监督学习和迁移学习中表现出色。DECREE\cite{feng2023detecting}提出了一种用于检测自监督学习中预训练图像编码器后门的创新方法，该方法无需依赖标签数据或下游分类器，且在多种数据集和攻击类型下表现出极高的检测准确率。
Backdoor Detection:
Backdoor detection primarily aims to identify existing backdoor attacks within a model. In this category, SEER \cite{zhu2024seer} proposes a backdoor detection algorithm for vision-language models, successfully detecting backdoors across various scenarios by jointly searching for target text and image triggers in the feature spaces of the image and text modalities. ASSET \cite{pan2023asset} introduces a robust backdoor data detection method across multiple deep learning paradigms by actively inducing behavioral differences between backdoor samples and clean samples in the model, demonstrating excellent performance in end-to-end supervised learning, self-supervised learning, and transfer learning. DECREE \cite{feng2023detecting} presents an innovative method for detecting backdoors in pre-trained image encoders used in self-supervised learning, achieving high detection accuracy across various datasets and attack types without relying on label data or downstream classifiers.

%后门缓解则侧重于通过各种技术手段来消除或减轻后门攻击的影响。Semantic Shield\cite{ishmam2024semantic}通过强制模型在训练时对齐图像区域与外部知识，以防御视觉-语言模型中的后门和数据中毒攻击，从而显著提高了模型的鲁棒性。MCLDef\cite{yue2023model}提出了一种基于模型对比学习的两阶段后门防御方法，通过收缩或破坏中毒数据在特征空间中的聚类，并将中毒数据的特征拉向其干净对应物，从而有效去除深度神经网络中的后门，同时在不显著降低模型准确性的情况下提升防御效果。DPoE\cite{liu2023shortcuts}提出了一种基于端到端集成的后门防御框架，通过结合浅层模型和主模型来捕捉和抑制后门触发器，从而有效应对各种显性和隐性的后门攻击，同时减轻噪声标签对模型性能的影响。此外，PSIM\cite{zhao2024defending}提出了一种基于参数高效微调的防御模块，通过利用样本的置信度来识别被后门攻击污染的样本，显著增强了模型抵御权重中毒后门攻击的能力，并在不影响模型准确性的情况下有效过滤出被污染的样本。CleanCLIP\cite{bansal2023cleanclip}通过在视觉和文本编码器上进行无监督微调，独立调整每种模态的表示，以削弱视觉-语言对比学习模型中由后门攻击引入的错误关联，从而有效减少后门攻击的影响，同时保持模型在正常数据上的性能。
Backdoor Mitigation:
Backdoor mitigation focuses on eliminating or reducing the impact of backdoor attacks through various technical means. Semantic Shield \cite{ishmam2024semantic} significantly enhances model robustness by aligning image regions with external knowledge during training to defend against backdoor and data poisoning attacks in vision-language models. MCLDef \cite{yue2023model} proposes a two-stage backdoor defense method based on model contrastive learning, effectively removing backdoors in deep neural networks by contracting or disrupting the clustering of poisoned data in the feature space and pulling the features of poisoned data toward their clean counterparts, while improving defense effectiveness without significantly reducing model accuracy. DPoE \cite{liu2023shortcuts} introduces an end-to-end integrated backdoor defense framework that captures and suppresses backdoor triggers by combining a shallow model and a main model, effectively countering various explicit and covert backdoor attacks while mitigating the impact of noisy labels on model performance. Additionally, PSIM \cite{zhao2024defending} proposes a defense module based on parameter-efficient fine-tuning, significantly enhancing the model's ability to resist weight poisoning backdoor attacks by identifying samples polluted by backdoor attacks using sample confidence, effectively filtering out polluted samples without affecting model accuracy. CleanCLIP \cite{bansal2023cleanclip} reduces the impact of backdoor attacks in vision-language contrastive learning models by independently adjusting the representations of each modality through unsupervised fine-tuning of the visual and text encoders, while maintaining model performance on clean data.

%尽管这些方法在应对后门攻击方面展现了显著效果，但在面对更隐蔽或复杂的后门攻击时，效果依然可能不够理想。某些方法还依赖于外部知识或置信度的设定，可能导致误判或性能瓶颈。此外，这些方法在实际应用中也增加了计算开销和训练时间。因此，未来的研究仍需在提升整体性能与效率的同时，着力解决隐蔽或复杂后门攻击的防御挑战。
Despite the significant effectiveness of these methods in countering backdoor attacks, their performance may still be insufficient when faced with more covert or complex backdoor attacks. Some methods also rely on external knowledge or confidence settings, which may lead to misjudgments or performance bottlenecks. Moreover, these methods increase computational overhead and training time in practical applications. Therefore, future research must continue to focus on addressing the challenges posed by covert or complex backdoor attacks while enhancing overall performance and efficiency.

\section{System model \& Preliminaries}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%symbols 表格
\begin{table}[!t]
{
\renewcommand{\arraystretch}{1.2}
\caption{Summary of Main Notation}
\label{notations}
\vspace{0em}
\centering
\begin{tabular}{| c || l |}
\hline
\textbf{Symbol} & \multicolumn{1}{c |}{\textbf{Description}}\\
    \hline
    $U_i,\mathcal{D}_i$ & The $i$-th UAV devices and its dataset\\
    \hline
    $n^{s_{m}, i},\mathcal{P}^i$ & The size and normalized distribution of $\mathcal{D}_i$\\
    \hline
    $\mathcal{P}^{Tar}, w$ & Target distribution and model of learning task $Job$\\
    \hline
    $T$ & The maximum communication rounds of $Job$\\
    \hline
    $C,K$ & The set and number of all peer communities\\
    \hline
    $S,M$ & The set and number of all colleague communities\\
    \hline
    %$L$ & The number of clients in each super node\\
    %\hline
    $c_k$ & The $k$-th peer community\\
    \hline
    $s_m$ & The $m$-th colleague community\\ 
    \hline
    $\Delta \widetilde{\omega}_{t}^{s_{m},i}$ & The weight updates from $U_i \in s_m$ in round $t$\\
    \hline
    $p^{s_{m}, i}$ & The aggregation coefficient for $U_i$ within $s_m$\\
    \hline
    $\Delta\widetilde{\omega}_{t}^{s_{m}}$ & The weight updates from $s_m$ in round $t$\\
    \hline
    $p^{s_{m}}$ & The aggregation coefficient for $\Delta\widetilde{\omega}_{t}^{s_{m}}$ from $s_m$\\
    \hline
    $E_k$ & The clipping bound of peer community $c_k$\\
    \hline
    $Sen_k$ & The maximum sensitivity of $c_k$\\
    \hline
    $\varepsilon_k,\delta_k$& The parameters of differential privacy \\
    \hline
    $L$& The minimum size of colleague community\\
    \hline
    $\alpha^{a,b}_{t}$& The cosine similarity between $a$ and $b$\\
    \hline
    $PI^{c_k \rightarrow i}_{t}$ & The positive interaction between $c_k$ and $U_i$\\
    \hline
    $NI^{c_k \rightarrow i}_{t}$& The positive interaction between $c_k$ and $U_i$\\
    \hline
    $\Upsilon^{c_k \rightarrow i}_{t}$& The peer reputation opinion vector of $c_k$\\
    \hline
    $\Phi^{c_k \rightarrow i}_{t}$& $c_k$'s credit evaluation towards $U_i$ in round $t$\\
    \hline
    $\Upsilon^{s_m}_{t}$ & The colleague reputation opinion vector towards $s_m$\\
    \hline
    $\Phi^{s_m \rightarrow i}_{t}$& The colleague credit evaluation for $\forall U_i \in s_m$\\
    \hline
    $\Upsilon^{i}_{t},\Phi^{i}_{t}$& The comprehensive reputation and credit of $U_i$\\
    \hline
    $\psi^{i}_t$& The fairness-guaranteed selection probability\\
    \hline
    \end{tabular}
        }
     \vspace{-1.5em}
     \end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Consider a scenario where a vision-language model is fine-tuned in a federated manner within an Internet of Robotic Things (IoRT) network. The system model comprises multiple robotic nodes and a central server. The robotic nodes perform local model fine-tuning, with global aggregation carried out through the central server. The following is a detailed description of the system model.

\textbf{Participants and Datasets:}
Assume the system model consists of $n$ robotic nodes, each denoted by $R_i$, where $i \in {1, 2, \ldots, n}$. Each robotic node $R_i$ possesses a local dataset $D_i = {x_{i,j}, y_{i,j}}{j=1}^{|D_i|}$, where $x{i,j}$ represents the input data, and $y_{i,j}$ represents the corresponding label data, with $|D_i|$ denoting the size of the dataset. The overall dataset is denoted by $D = \bigcup_{i=1}^{n} D_i$.

\textbf{LoRA:}
In the Low-Rank Adaptation (LoRA) method, each parameter matrix $W$ is represented as:

where $W_0$ is the original weight matrix of the model, $\Delta W$ is the low-rank increment matrix, and $B$ and $A$ are the low-rank decomposition matrices, with $r$ representing the rank of the matrices. The original model parameters are denoted by $\Theta = \sum {W_0}$, while the introduced incremental parameters are denoted by $\theta = \sum {W}$. During federated fine-tuning, only the incremental parameters $\theta$ are updated, while the original parameters $\Theta$ remain unchanged.

\textbf{Local Model Training:}
At the start of each federated training round, the central server sends the global incremental parameters $\theta_g^{t-1}$ to each robotic node $R_i$. Upon receiving the incremental parameters $\theta_g^{t-1}$, each node fine-tunes its local model based on its local dataset $D_i$.
Assume the number of local training epochs is $E_l$, the batch size is $b$, and the learning rate is $\eta$. For each epoch, node $R_i$ first randomly samples a batch of data $D_i^b \sim Sample(D_i, b)$ from its local dataset $D_i$, then updates the incremental parameters using optimization algorithms such as stochastic gradient descent:

where $(\theta_i^{t})^e$ denotes the parameters after the $e$-th fine-tuning epoch, with the initialization $(\theta_i^{t})^0 = \theta_g^{t-1}$. The objective of local training is to minimize the loss function $f(D_i^b, (\theta_i^{t})^e)$:

where $\Theta$ represents the original model parameters, and $\ell$ denotes the loss function (e.g., cross-entropy loss). After completing local training, the node computes its incremental parameter update $\triangledown \theta_i^t = (\theta_i^t)^{E_l} - (\theta_i^t)^0$ and uploads it to the central server.

\textbf{Federated Aggregation:}
At the end of each round of local training, the central server collects the local updates $\triangledown \theta_i^t$ from each robotic node $R_i$ and aggregates them using the appropriate aggregation algorithm to update the global incremental parameters $\theta_g^t$, computed as:

where $Aggregate$ represents the aggregation algorithm employed by the central server, such as FedAvg, Krum, Bulyan, or RoseAgg. The updated global incremental parameters $\theta_g^t$ will be disseminated to all robotic nodes before the next round of training for further local fine-tuning.

\section{Threat model}
%Attacker's Goal：假设在对视觉语言模型进行联邦微调的场景中，存在一个或者多个由攻击者操纵的恶意节点。其目标是对全局模型$\theta_g$ 进行后门植入，使得模型能够将任何带有预定义触发器 $\delta$ 的测试输入分类为目标类别，同时尽可能保持对干净输入的分类准确性。
\textbf{Attacker's goal:} Consider a scenario where one or more malicious nodes controlled by an attacker are present during the FFT process of a vision-language model. The attacker's objective is to implant a backdoor into the global model $\theta_g$, enabling the model to classify any test input containing a predefined trigger $\delta$ as the target class, while simultaneously maintaining the classification accuracy on clean inputs to the greatest extent possible.

%Attacker's Ability: 在联邦学习场景中，假设攻击者对恶意节点的本地训练过程具有完全的控制权。这意味着攻击者可以任意操控本地数据、触发器模式、优化策略和本地更新过程。此外，攻击者可以采用“智能”攻击策略，这种策略的特点是攻击效果不会在单轮训练中显现，而是通过多轮训练的累积效应逐步将后门植入全局模型中。这种场景在实际中是可能存在的，因为中央服务器只能接收到客户端上传的训练后模型更新，而无法知晓每个客户端的具体训练过程。然而，攻击者无法对中央服务器的操作进行干扰，例如，无法更改全局模型的聚合规则或篡改其他良性客户端的模型更新。
\textbf{Attacker's ability:} In the distributed training scenario as FL, it is assumed that the attacker has full control over the local training process on the malicious nodes. This means the attacker can manipulate local data, trigger patterns, optimization strategies, and the local update process at will. Furthermore, the attacker may employ a ``stealthy" attack strategy, characterized by the backdoor being gradually implanted into the global model through the cumulative effect of multiple training rounds, rather than becoming apparent in a single training round. Such a scenario is plausible in practice because the central server only receives the model updates uploaded by robot nodes after training, without knowledge of the specific training process for each client. However, the attacker cannot interfere with the central server's operations, such as altering the global model aggregation rules or tampering with the model updates from other benign nodes.

\section{Method}

\subsection{Single}

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/fig2.pdf}
    \caption{Benign and Malicious Nodes Before Aggregation}
    \label{fig2:single}
\end{figure}

\subsection{Multi}

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/fig3.jpg}
    \caption{Benign and Malicious Nodes Before Aggregation}
    \label{fig3:multi}
\end{figure}

\begin{figure}[!t]
    \centering
    \includegraphics[width=\linewidth]{figures/fig4.jpg}
    \caption{Benign and Malicious Nodes Before Aggregation}
    \label{fig4:before-agg}
\end{figure}

@@

@@

\section{System model \& Preliminaries}
%%symbols 表格
\begin{table}[!t]
{
\renewcommand{\arraystretch}{1.2}
\caption{Summary of Main Notation}
\label{notations}
\vspace{0em}
\centering
\begin{tabular}{| c || l |}
\hline
\textbf{Symbol} & \multicolumn{1}{c |}{\textbf{Description}}\\
    \hline
    $U_i,\mathcal{D}_i$ & The $i$-th UAV devices and its dataset\\
    \hline
    $n^{s_{m}, i},\mathcal{P}^i$ & The size and normalized distribution of $\mathcal{D}_i$\\
    \hline
    $\mathcal{P}^{Tar}, w$ & Target distribution and model of learning task $Job$\\
    \hline
    $T$ & The maximum communication rounds of $Job$\\
    \hline
    $C,K$ & The set and number of all peer communities\\
    \hline
    $S,M$ & The set and number of all colleague communities\\
    \hline
    %$L$ & The number of clients in each super node\\
    %\hline
    $c_k$ & The $k$-th peer community\\
    \hline
    $s_m$ & The $m$-th colleague community\\ 
    \hline
    $\Delta \widetilde{\omega}_{t}^{s_{m},i}$ & The weight updates from $U_i \in s_m$ in round $t$\\
    \hline
    $p^{s_{m}, i}$ & The aggregation coefficient for $U_i$ within $s_m$\\
    \hline
    $\Delta\widetilde{\omega}_{t}^{s_{m}}$ & The weight updates from $s_m$ in round $t$\\
    \hline
    $p^{s_{m}}$ & The aggregation coefficient for $\Delta\widetilde{\omega}_{t}^{s_{m}}$ from $s_m$\\
    \hline
    $E_k$ & The clipping bound of peer community $c_k$\\
    \hline
    $Sen_k$ & The maximum sensitivity of $c_k$\\
    \hline
    $\varepsilon_k,\delta_k$& The parameters of differential privacy \\
    \hline
    $L$& The minimum size of colleague community\\
    \hline
    $\alpha^{a,b}_{t}$& The cosine similarity between $a$ and $b$\\
    \hline
    $PI^{c_k \rightarrow i}_{t}$ & The positive interaction between $c_k$ and $U_i$\\
    \hline
    $NI^{c_k \rightarrow i}_{t}$& The positive interaction between $c_k$ and $U_i$\\
    \hline
    $\Upsilon^{c_k \rightarrow i}_{t}$& The peer reputation opinion vector of $c_k$\\
    \hline
    $\Phi^{c_k \rightarrow i}_{t}$& $c_k$'s credit evaluation towards $U_i$ in round $t$\\
    \hline
    $\Upsilon^{s_m}_{t}$ & The colleague reputation opinion vector towards $s_m$\\
    \hline
    $\Phi^{s_m \rightarrow i}_{t}$& The colleague credit evaluation for $\forall U_i \in s_m$\\
    \hline
    $\Upsilon^{i}_{t},\Phi^{i}_{t}$& The comprehensive reputation and credit of $U_i$\\
    \hline
    $\psi^{i}_t$& The fairness-guaranteed selection probability\\
    \hline
    \end{tabular}
        }
     \vspace{-1.5em}
     \end{table}

@@

% \begin{figure}[!t]
% \centering
% \includegraphics[width=\linewidth]{figures/fig2.PNG}
% \caption{Stratified community construction in a heterogeneous environment.}
% \label{fig2:framework}
% \end{figure}

@@

% \begin{figure}[!t]
% \centering
% \includegraphics[width=\linewidth]{figures/fig3.PNG}
% \caption{The workflow of MR-FFL.}
% \label{fig3:workflow}
% \end{figure}

@@

\begin{algorithm}
\caption{Single}
\label{alg:malicious-node-detection-history}
\begin{algorithmic}[1]
\State \textit{\textbf{Input:} Global model $\{\theta_g^{t-T}, \dots, \theta_g^{t-1}\}$, historical updates for each nodes $\{\triangledown \theta_{i}^{t-T}, \dots, \triangledown \theta_{i}^{t-1}\}$, threshold $\varepsilon$.}
\State \textit{\textbf{Output:} Sets of normal/malicious nodes $U_{nor}$ and $U_{mal}$.}

\State {\color{CadetBlue}/*\textbf{Step 1: historical database construction*/}}
\For {Client $i \in $}
    \State Record flattened gradients $\bar{\theta_i^t}$ and models $\theta_g^{t-1};$
    \State Establish a ray with Eq. () for each of the $t'$ most recent gradients, $t' = \max\{t-T+1, 1\}, \ldots, t$;
\EndFor

\State \textbf{Step 2: Obtain the Purpose Intention}
\For {Client $i$}
    \State Initialize center $O_{i,0}$ and radius $r_{i,0}$
    \While {Not Converged}
        \State Update center $O_{i,k+1}$ and radius $r_{i,k+1}$
    \EndWhile
    \State Calculate confidence $cre_i = \frac{1}{r_i + \rho}$
\EndFor

\State \textbf{Step 3: Abnormal Detection}
\For {Client $i$}
    \For {Each point $O_j \neq O_i$}
        \State Calculate distance $\|O_i - O_j\|$
    \EndFor
    \State Determine $q$-distance $\widetilde{dis_i^q}$ for $O_i$
    \State Find neighbors $Nei_i$ within $\widetilde{dis_i^q}$
    \State Compute reachability distance $rea_{i,j} = \max\{\widetilde{dis_i^q}, \|O_i - O_j\|\}$ for $O_j \in Nei_i$
    \State Calculate local reachability density $lrd_i = \frac{|Nei_i|}{\sum_{j \in Nei_i} rea_{i,j}}$
\EndFor
% \For {Client $i$}
%     \State Compute $lof_i = \frac{\sum_{j \in Nei_i} \frac{lrd_j}{lrd_i}}{|Nei_i|}$
%     \If {$lof_i > 1$}
%         \State Mark client $i$ as malicious, add to $U_{mal}$
%     \Else
%         \State Mark client $i$ as normal, add to $U_{nor}$
%     \EndIf
% \EndFor
\State Calculate $lof_i = \frac{\sum_{j \in Nei_i} \frac{lrd_j}{lrd_i}}{|Nei_i|}$ for each client $i$
% \State $U_{mal} = \{i \,|\, lof_i > 1\}$
% \State $U_{nor} = \{i \,|\, lof_i \leq 1\}$
\State $U_{mal}, U_{nor} = \{i \,|\, lof_i > 1\}, \{i \,|\, lof_i \leq 1\}$

\State \textbf{Step 4: Gradients Aggregation}
\For {Client $i\in U_{nor}$}
        \State Compute adjusted confidence $cre_i' = \tanh(cre_i)$
        \State Normalize weights $w_i = \frac{cre_i'}{\sum_{i \in U_{nor}} cre_i'}$
\EndFor

\State \textbf{Output:} Aggregated global model update $\theta_g^t = \theta_g^{t-1} + \sum_{i \in U_{nor}} w_i \cdot \triangledown \theta_i^t$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Malicious Node Detection via DCT}
\label{alg:malicious-node-detection}
\begin{algorithmic}[1]
\State \textbf{Input:} $d$, $n$, $(\triangledown \theta_0^t, \triangledown \theta_1^t, \ldots, \triangledown \theta_n^t) \in \mathbb{R}^{n \times d}$, $m$ \Comment{$d$ is the dimension of each node update; $n$ is the number of the nodes participating during each round; $(\triangledown \theta_0^t, \triangledown \theta_1^t, \ldots, \triangledown \theta_n^t) \in \mathbb{R}^{n \times d}$is the local updates from nodes during the $t$-th round; $m$ is the length of low-frequency components}
\State \textbf{Output:} $U_{nor}$, $U_{mal}$ \Comment{benign nodes, malicious nodes}

\State Step 1: Frequency Domain Transformation
\For {Node $R_i$}
    \State $\bar{\theta_i^t} \gets Flatten(\triangledown \theta_i^t)$
    \State $G_i^t = Trunc(DCT(\bar{\theta_i^t}), m)$ \Comment{Compute DCT and retain only the top $m$ low-frequency components}
\EndFor

\State

\State Step 2: Clean Ingredient Extraction
\State $H^t \gets (G_0^t, G_1^t, \ldots, G_{n}^t) \in \mathbb{R}^{d \times m}$ \Comment{Stacking to form Matrix $H^t$.}

\State $\hat{H}^t \gets (H^t)^{T} H^t$
\State $\lambda_{max}^t, \xi_{max}^t \gets \text{eig}(\hat{H}^t)$ \Comment{Calculating the maximum singular value and its corresponding eigenvector}
\State $\tilde{G}^t \gets \frac{H^t \xi_{max}^t}{\sqrt{\lambda_{max}^t}}$ \Comment{The clean ingredient}

\State

\State Step 3: Kullback-Leibler Divergence Calculation
\State $P_i^t \gets Softmax(G_i^t), \, Q^t \gets Softmax(\tilde{G}^t)$
\For {Node $R_i$}
    \State $Ki \gets \sum_{k=0}^{m-1} P_i^t[k] \log \left( \frac{P_i^t[k]}{Q^t[k]} \right)$
\EndFor
\State $S \gets \{KL_1, KL_2, \ldots, KL_n\}$ \Comment{The distribution differences calculated by KL divergence.}

\State

\State Step 4: Malicious Node Detection
\State $\{C_1, C_2, \ldots, C_K\} \gets \text{HDBSCAN}(S)$
\State $U_{nor} \gets C_{\text{max}}$, $U_{mal} \gets \{C_i \, | \, i \neq \text{max}\}$

\end{algorithmic}
\end{algorithm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\bibliographystyle{IEEEtran}
\bibliography{ref,attack_defense}

@@

% \begin{figure}[!t]
% \centering
% \includegraphics[width=\linewidth]{figures/fig4-history.jpg}
% \caption{Combining multiple rounds of historical identification to identify malicious nodes
% }
% \label{fig4:history}
% \end{figure}

% \begin{figure}[!t]
% \centering
% \includegraphics[width=\linewidth]{img/SecFFT-场景.pdf}
% \caption{Combining multiple rounds of historical identification to identify malicious nodes
% }
% \label{fig4:history}
% \end{figure}

% \begin{figure}[!t]
% \centering
% \includegraphics[width=\linewidth]{img/SecFFT-算法1.pdf}
% \caption{Combining multiple rounds of historical identification to identify malicious nodes
% }
% \label{fig4:history}
% \end{figure}
